---
title: "HP Filter Quarterly Data"
output: html_notebook
---


```{r}
library(mFilter)
library(dplyr)
library(ggplot2)
library(dygraphs)
library(xts)
library(readxl)
library(quantmod)
library(neverhpfilter)
```

```{r}
options(prompt="R> ", digits=3, scipen=999)
```

```{r}
library(readxl)
mydata <- read_excel("Output gap_quarterly.xlsx")
mydata
```

```{r}
library(tidyr)
```

```{r}
mydata$GRDP = as.numeric(mydata$GRDP)
```

```{r}
library(reshape2)
```

```{r}
data_w <- dcast(mydata, Time ~ Province, value.var = "GRDP")
```

```{r}
data_w
```

# define data frame as time series
```{r}
mydatats <- ts(data_w[,2:36], start =c(2010,1), frequency = 4)
```

```{r}
mydatats
```

```{r}
str(mydatats)
```

```{r}
View(mydatats)
```

# log transform
```{r}
lngdp <- 100*log(mydatats)
```

```{r}
View(lngdp)
```

# Plot Aceh GRDP (Level and log)
```{r}
par(mfrow=c(1,2))
plot(mydatats[,1], xlab="", ylab="", main="Aceh GDP level")
plot(lngdp[,1], xlab="", ylab="", main="Aceh GDP logs")
par(mfrow=c(1,1))
```


```{r}
library(dygraphs)
```

```{r}
dygraph(lngdp[,1]) %>% 
  dyRangeSelector()
```
# Apply HP Filter to series column 1 to 34 (Aceh to Yogyakarta) in loop


# alternative 1: create time column manually
```{r}
year = rep(2010:2019, each =4)
year
quarter = rep(c("q1", "q2", "q3", "q4"), 10)
quarter
time <- paste(year, quarter, sep = "_")
time
```

# Add rows Q1 and Q2 2020 
```{r}
hp <- data.frame(time)
```

```{r}
hp <- rbind(hp, c("2020_q1")) %>% 
  rbind(c("2020_q2"))
hp 
```

# join time column to lngdp
```{r}
lngdp1 <- data.frame(hp, lngdp)
```


# alternative 2: extract time column from data_w
```{r}
hp1 <- data.frame(data_w$Time)
hp1
```

#Extract time column from lngdp dataframe. We use it for joining actual, trendHP, cycleHP with hp  
```{r}
time1 <- data.frame(time=lngdp1[,1])
time1
```

# hp filter for all provinces in loops
```{r}
for (i in 2:36) {
  hpfil <- hpfilter(lngdp1[,i], freq = 1600)
actual <- hpfil[["x"]]
trendHP <- hpfil[["trend"]]
cycleHP <- actual - trendHP

colnames(actual) <- c(paste(colnames(lngdp1)[i],"actual",sep = "_"))
colnames(trendHP) <- c(paste(colnames(lngdp1)[i],"trendHP",sep = "_"))
colnames(cycleHP) <- c(paste(colnames(lngdp1)[i],"cycleHP",sep = "_"))

actual_trend_cycle <- data.frame(actual, trendHP, cycleHP, time1)

hp <- left_join(hp, actual_trend_cycle, by = "time")
}
```

# check the computed hp filter results for all provinces combined
```{r}
View(hp)
```

#save the hp dataframe to csv
```{r}
write.table(hp, "outputgap.csv", sep = ";", col.names = NA,
            qmethod = "double")
```

#extract only cycle column
```{r}
hpcycle <- select(hp, -contains("actual")) %>% 
  select(-contains("trendHP"))
```


```{r}
long_hp <- pivot_longer(hpcycle, -c(time), values_to = "output_gap", names_to = "province" )
```

```{r}
long_hp <- long_hp %>% mutate(province = as.character(gsub("_cycleHP", "", province)))
```

```{r}
graph <- long_hp %>% 
  ggplot(aes(x = time, y = output_gap, color= province)) +
geom_line(size=0.5) +
labs(subtitle = "Output Gap, 2012-2019",
x = "Time",
y = "Output gap", color="") + theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())
graph
```

```{r}
q2_2020hp <- filter(long_hp, time == "2020_q2")
```


```{r}
outputgap <- ggplot(q2_2020hp, 
       aes(x = reorder(province, -output_gap), 
           y = output_gap, 
           fill = ifelse(province == "Indonesia", "Highlighted", "red") )) + 
  geom_bar(stat = "identity") + 
   theme_minimal() +
labs(subtitle = "",
x = "",
y = "Output gap in Q2-2020 (%)") +
theme(text=element_text( family="Palatino")) +
  ## drop legend and Y-axis title
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "none", axis.title.y = element_blank()) +
  coord_flip(expand = FALSE) ########## SMALL UPDATE
outputgap
ggsave("outputgap.png", width = 9, height = 5)
```

